---
title: "DDS Project 1 - Beer Data Analysis"
author: "Andrew Leppla and Anish Patel"
date: "3/7/2020"
output: html_document
---
##Introduction

This project analyzes beer and brewery data.  Exploratory data analysis is conducted on the data set to determine which states in the USA are dominant in brewing beer.  The data set contains missing values and the researchers analysis details the best way to deal with these missing values.  The researchers will explore relationships between ABV and IBU in addition to geographical trends.  The researchers will build models to attempt in predicting given beer styles from ABV and IBUs of a beer and investigate other interesting relationships in the beer data set.

knitr::opts_chunk$set(echo = TRUE)
```

```{r import, include=FALSE}
setwd("~/SMU-Doing-Data-Science-Project1")
Beers = read.csv("Beers.csv",header=T)
Breweries = read.csv("Breweries.csv",header=T)
states = read.csv("states.csv")

#Download from Kaggle, file too large for GitHub repo 
#(https://www.kaggle.com/rdoume/beerreviews)
beer_reviews = read.csv("beer_reviews.csv") 
```

```{r libraries, include=FALSE}
library(dplyr)
library(tidyverse)
library(maps)
library(mapproj)
library(ggmap)
library(ggplot2)
library(mapdata)
library(openintro)
library(plotly)
library(billboarder)
library(vctrs)
library(stringr)
library(class) #knn function
library(caret) #confusionMatrix function
```
## Introduction

This project analyzes beer and brewery data.  Exploratory data analysis is conducted on the data set to determine which states in the USA are dominant in brewing beer.  The data set contains missing values and the researchers analysis details the best way to deal with these missing values.  The researchers will explore relationships between ABV and IBU in addition to geographical trends.  The researchers will build models to attempt in predicting given beer styles from ABV and IBUs of a beer and investigate other interesting relationships in the beer data set relating to ABV, IBU, and beer ratings.  



## 1. How many breweries are present in each state?

Colorado and California have the most breweries.  Each US region has 1-2 states with more breweries as well:

- Northwest: Oregon and Washington

- South: Texas

- Midwest: Michigan

- Northeast: Pennsylvania and Massachusettes

```{r Breweries by State, echo=FALSE}
#Table of Number(No.) of Breweries by State in decreasing order
State_Breweries = Breweries %>% group_by(State) %>% summarize(No.Breweries = n()) %>% arrange(-No.Breweries)
State_Breweries_df=as.data.frame(State_Breweries)

#Plot of No. Breweries by State in alphabetical order
State_Breweries_df %>% ggplot(aes(State,No.Breweries,fill=State)) + geom_col() + ggtitle("Number of Breweries by State - Alphabetical") + ylab("# Breweries") + theme(axis.text.x=element_blank()) 

#Plot of No. Breweries by State in decreasing order
State_Breweries_df$State1=factor(State_Breweries_df$State,levels=State_Breweries_df$State[order(-State_Breweries_df$No.Breweries)]) 
State_Breweries_df %>% ggplot(aes(State1,No.Breweries,fill=State1)) + geom_col() + ggtitle("Number of Breweries by State - Decreasing") + ylab("# Breweries") + xlab("State") + theme(axis.text.x=element_blank())

#Heatmap of Breweries by State
brew_map = ggplot(data = states) + 
  geom_polygon(aes(x = long, y = lat, fill = n, group = group), color = "white") + 
  coord_fixed(1.3) +
  scale_fill_gradientn(colours = rev(heat.colors(10)),na.value = "grey90",
  guide = guide_colourbar(barwidth = 5, barheight = 0.4,
  title.position = "top")) + 
  labs(title = "Count of Breweries by State (Continental US)")
  
ggplotly(brew_map)
```

## 2. Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file. 
```{r Merge Beers and Breweries, echo=FALSE}
#Inspect the data for merging consistency

#summary(Breweries$Brew_ID) #558 breweries
#summary(Beers$Brewery_id) #558 breweries

#Outer Merge
Beer = merge(Breweries,Beers,by.x="Brew_ID",by.y="Brewery_id",all = TRUE,suffixes=c(".Brewery",".Beer"))

head(Beer,n=6)
tail(Beer,n=6)
```

## 3. Address the missing values in each column.

There are 5 missing values for Style, 62 missing values for ABV, and 1,005 missing values for IBU.  

Beer Style is a grouping/classification for beers that's been established by brewers based on brewing traditions and their domain expertise.  Beers within an established Beer Style tend to have more similar alcohol content and bitterness (low within group variation), whereas beers that differ in Style tend to have less similar ABV and IBUs (high between group variation).  Based on this knowledge, median ABV and IBU values were calculated for all 100 Beer Styles.  Missing ABV and IBU values were generally addressed by replacing NA's with the matching Beer Style's median values.                

2 of the 5 missing Style values were imputed based on the individual beer names.  "OktoberFiesta" had a beer name, ABV, and IBU that were consistent with the other Oktoberfest style beers.  Similarly, "Kilt Lifter Scottish-Style Ale" was consistent with the Scottish Ales style beers.  The remaining 3 beers with missing Style values did not have enough information to classify their Style.  For those 3, Style was left blank, and the missing IBU and ABV values were set to the overall median values of the entire data set.

55 of the 1,005 missing IBU values could not be imputed this way because those Styles were missing all IBU values.  These 55 beers were associated with 10 unique Styles that had no IBU values.  Ciders, Meads, Shandies, and Rauchbiers (smoked beers) styles are all typically made with no hops or bittering of any kind.  These missing IBU values were set to 0 based on this domain knowledge.  There were 2 exceptions where the product names indicated there may be some hops added, contrary to the style conventions: Cider "Nunica Pine" and Mead "Nectar of Hops".  For these 2, the missing IBU values were set to the overall median IBU value for the entire data set.            

Finally, the remaining 11 missing IBU values were set to the overall median IBU value for the entire data set.  

All further analyses were done by both excluding the missing values and by using the imputed values. 

```{r Merge and Missing Values, include=FALSE}
summary(Beer) #62 NA ABV values (2-3%), 1005 NA IBU values (about 40%)

#Some Styles have European special characters that are read in differently/inconsistently.
#Not reproducible, replace with latin alphabetical characters
levels(Beer$Style)[c(34,69,74)] #Styles with special characters
levels(Beer$Style)[34]="Biere de Garde"
levels(Beer$Style)[69]="Kolsch"
levels(Beer$Style)[74]="Marzen / Oktoberfest"

#Beer %>% filter(is.na(ABV)) #Some Styles are blank ""
Beer %>% filter(Style=='') #5 missing beer Styles  

#Replace 2 of the missing Styles based on Beer Name 
Beer$Style[Beer$Name.Beer=="OktoberFiesta"]="Marzen / Oktoberfest"
Beer$Style[Beer$Name.Beer=="Kilt Lifter Scottish-Style Ale"]="Scottish Ale"
Beer$Style[Beer$Style=='']=NA

#Mean ABV by Style
Style_ABV = Beer %>% filter(!is.na(ABV)) %>% group_by(Style) %>% summarize(median.ABV=median(ABV),count=n()) %>% arrange(Style)
Style_ABV_df=as.data.frame(Style_ABV)
dim(Style_ABV_df) # 99/100 Styles

#Replace the NA ABVs with the median ABV by Style
Beer = merge(Beer,Style_ABV,by="Style",all=T)
Beer$ABV1=ifelse(is.na(Beer$ABV),Beer$median.ABV,Beer$ABV)
summary(Beer$ABV1) #3 remaining NAs

#For remaining 3 missing Styles, replace with the overall median ABV
summary(Beer$ABV1) # 5.6% overall median ABV
Beer$ABV1=ifelse(is.na(Beer$ABV1),summary(Beer$ABV)[3],Beer$ABV1)

#Check that all ABV NAs have been replaced
summary(Beer$ABV1) #No NA's
length(Beer$ABV1) #2410 rows


#Median IBU by Style
Style_IBU = Beer %>% filter(!is.na(IBU)) %>% group_by(Style) %>% summarize(median.IBU=median(IBU),count=n()) %>% arrange(Style)
Style_IBU_df=as.data.frame(Style_IBU) 
dim(Style_IBU_df) # 90/100 Styles, missing 10 Styles

#Replace the NA IBUs with the median IBU by Style
Beer = merge(Beer,Style_IBU,by="Style",all=T)
Beer$IBU1=ifelse(is.na(Beer$IBU),Beer$median.IBU,Beer$IBU)

#Remaining NA IBUs
summary(Beer$IBU1) #55/1005 beers remaining with IBU NA values
#Ciders, Meads, Shandies, and Rauchbiers (smoked beers) typically have IBU=0
Beer$IBU1=ifelse(Beer$Style=="Cider" | Beer$Style=="Mead" | Beer$Style=="Shandy" | Beer$Style=="Rauchbier",0,Beer$IBU1) #Replace NAs with 0 for these Styles
summary(Beer$IBU1) #8/1005 beers remaining with IBU NA values

#Cider "Nunica Pine" and Mead "Nectar of Hops" probably have nonzero IBU
#Set to overall median IBU
Beer$IBU1[Beer$Name.Beer=="Nunica Pine" | Beer$Name.Beer=="Nectar of Hops"]=summary(Beer$IBU)[3]

#Replace the remaining 8 NA IBUs with overall median IBU
Beer$IBU1=ifelse(is.na(Beer$IBU1),summary(Beer$IBU)[3],Beer$IBU1)

#Check that all NA IBUs have been replaced
summary(Beer$IBU1) #No NA values
length(Beer$IBU1) #2410 rows

```


```{r Median ABV and IBU by State - Excluded, include=FALSE}
#Median ABV by State with ABV NAs
State_ABV = Beer %>% filter(!is.na(ABV)) %>% group_by(State) %>% summarize(Median.ABV=median(ABV)*100) %>% arrange(-Median.ABV)
State_ABV_df=as.data.frame(State_ABV)

#Plot of ABV by State in alphabetic order
State_ABV_df %>% ggplot(aes(State,Median.ABV,fill=State)) + geom_col() + ggtitle("Median Beer ABV by State - Excludes Missing Values") + theme(legend.position = "none") 

#Plot of ABV by State in decreasing order
State_ABV_df$State1=factor(State_ABV_df$State,levels=State_ABV_df$State[order(-State_ABV_df$Median.ABV)]) 
State_ABV_df %>% ggplot(aes(State1,Median.ABV,fill=State1)) + geom_col() + ggtitle("Median Beer ABV by State - Excludes Missing Values") + xlab("State") + theme(legend.position = "none") 

#Top 15 States for Highest Median ABV
head(State_ABV_df,n=15) %>% ggplot(aes(State1,Median.ABV,fill=State1)) + geom_col() + ggtitle("Excludes Missing Values") + xlab("State") + theme(legend.position = "none") 


#Median IBU by State
State_IBU = Beer %>% filter(!is.na(IBU)) %>% group_by(State) %>% summarize(Median.IBU=median(IBU)) %>% arrange(-Median.IBU)
State_IBU_df=as.data.frame(State_IBU)

#Plot of IBU by State in alphabetic order
State_IBU_df %>% ggplot(aes(State,Median.IBU,fill=State)) + geom_col() + ggtitle("Median Beer IBU by State- Excludes Missing Values") + theme(legend.position = "none") 

#Plot of ABV by State in decreasing order
State_IBU_df$State1=factor(State_IBU_df$State,levels=State_IBU_df$State[order(-State_IBU_df$Median.IBU)]) 
State_IBU_df %>% ggplot(aes(State1,Median.IBU,fill=State1)) + geom_col() + ggtitle("Median Beer IBU by State - Excludes Missing Values") + xlab("State") + theme(legend.position = "none") 

#Top 15 States for Highest Median IBUs
head(State_IBU_df,n=15) %>% ggplot(aes(State1,Median.IBU,fill=State1)) + geom_col() + ggtitle("Excludes Missing Values") + xlab("State") + theme(legend.position = "none") 

```


```{r Median ABV and IBU by State - Imputed, include=FALSE}
#Overall medians
summary(Beer$ABV1)[3]*100 # 5.6% ABV
summary(Beer$IBU1)[3] # 32 IBUs

#Median ABV by State 
State_ABV1 = Beer %>% group_by(State) %>% summarize(Median.ABV=median(ABV1)*100) %>% arrange(-Median.ABV)
State_ABV1_df=as.data.frame(State_ABV1)
State_ABV1_df

#Plot of ABV by State in alphabetic order
State_ABV1_df %>% ggplot(aes(State,Median.ABV,fill=State)) + geom_col() + ggtitle("Median Beer ABV by State - Includes Predicted Values for Missing Data") + theme(legend.position = "none") 

#Plot of ABV by State in decreasing order
State_ABV1_df$State2=factor(State_ABV1_df$State,levels=State_ABV1_df$State[order(-State_ABV1_df$Median.ABV)]) 
State_ABV1_df %>% ggplot(aes(State2,Median.ABV,fill=State2)) + geom_col() + ggtitle("Median Beer ABV by State - Includes Predicted Values for Missing Data") + xlab("State") + theme(legend.position = "none") 

#Top 15 States for Highest Median ABV
head(State_ABV1_df,n=15) %>% ggplot(aes(State2,Median.ABV,fill=State2)) + geom_col() + ggtitle("Includes Predicted Values for Missing Data") + xlab("State") + theme(legend.position = "none") 


#Median IBU by State  
State_IBU1 = Beer %>% group_by(State) %>% summarize(Median.IBU=median(IBU1)) %>% arrange(-Median.IBU)
State_IBU1_df=as.data.frame(State_IBU1)
State_IBU1_df

#Plot of IBU by State in alphabetic order
State_IBU1_df %>% ggplot(aes(State,Median.IBU,fill=State)) + geom_col() + ggtitle("Median Beer IBU by State - Includes Predicted Values for Missing Data") + theme(legend.position = "none") 

#Plot of IBU by State in decreasing order
State_IBU1_df$State2=factor(State_IBU1_df$State,levels=State_IBU1_df$State[order(-State_IBU1_df$Median.IBU)])
State_IBU1_df %>% ggplot(aes(State2,Median.IBU,fill=State2)) + geom_col() + ggtitle("Median Beer IBU by State - Includes Predicted Values for Missing Data") + xlab("State") + theme(legend.position = "none") 

#Top 15 States for Highest Median IBUs
head(State_IBU1_df,n=15) %>% ggplot(aes(State2,Median.IBU,fill=State2)) + geom_col() + ggtitle("Includes Predicted Values for Missing Data") + xlab("State") + theme(legend.position = "none") 

```

## 4. Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare. 

Utah has the lowest median ABV likely due to its strcit state alcohol laws and regulations.  Kentucky and Washington DC have the highest median ABV.  

Replacing missing values for ABV does not really change the trend.  This is not surprising since we're only missing 2-3% of the data.  Delaware does get a notable bump in the trend.     

Replacing missing values for IBU makes a big difference!  There are big swings in the trends after imputing 40% of the data based on beer styles.  The Northeastern states have the most notable shifts: Maine, New Hampshire, and Vermont.  Some of the lower values for NH are driven by missing data for sour beers which have very low IBUs.  The higher values for VT are driven by hoppier Pale Ales and IPAs that may be part of the trending "Juicy/Hazy" New England style IPAs.  


```{r, echo=FALSE}

ABV = merge(State_ABV_df,State_ABV1_df,by="State",suffixes=c(".Excl",".Impt"))

ABV %>% ggplot(aes(State1,Median.ABV.Excl,fill=State1)) + geom_col() + ggtitle("ABV by State - Excludes Missing Data") + xlab("State") + theme(axis.text.x=element_blank()) + ylab("Median ABV") + coord_cartesian(ylim=c(4,6.5))

ABV %>% ggplot(aes(State1,Median.ABV.Impt,fill=State1)) + geom_col() + ggtitle("ABV by State - Predicted Values for Missing Data") + xlab("State") + theme(axis.text.x=element_blank()) + ylab("Median ABV") + coord_cartesian(ylim=c(4,6.5))


IBU = merge(State_IBU_df,State_IBU1_df,by="State",suffixes=c(".Excl",".Impt"))

IBU %>% ggplot(aes(State1,Median.IBU.Excl,fill=State1)) + geom_col() + ggtitle("IBU by State - Excludes Missing Data") + xlab("State") + theme(axis.text.x=element_blank()) + ylab("Median IBU") 

IBU %>% ggplot(aes(State1,Median.IBU.Impt,fill=State1)) + geom_col() + ggtitle("IBU by State - Predicted Values for Missing Data") + xlab("State") + theme(axis.text.x=element_blank()) + ylab("Median IBU")
```

## 5. Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?

Colorado has the maximum alcoholic (ABV) beer of 12.8%, "Lee Hill Series Vol. 5 - Belgian Style Quadrupel Ale" by Upslope Brewing.

Oregon has the most bitter (IBU) beer of 138 IBU, "Bitter Bitch Imperial IPA" by Astoria Brewing Company.

```{r Max ABV and IBU, echo=FALSE}
Beer[Beer$ABV1==max(Beer$ABV1),] #CO
Beer[Beer$IBU1==max(Beer$IBU1),] #OR
```

## 6. Comment on the summary statistics and distribution of the ABV variable

ABV values range from 0.1% to 12.8% with a median of 5.6% and a mean of 6.0%.  The higher mean vs. the median indicates the distribution is right-skewed, and the histogram plot visually confirms this.  This right-skew may indicate a shift in the beer market towards "bigger" high alcohol beers.     

```{r Summary and Distribution, ech=FALSE}
summary(Beer$ABV1)

Beer %>% ggplot(aes(ABV1*100)) + geom_histogram(fill="darkblue",color="black") + xlab("% Alcohol by Volume (%ABV)") + ggtitle("Distribution of Beer %ABV, Right-Skewed")
```

## 7. Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.  Make your best judgment of a relationship and EXPLAIN your answer.

There appears to be an approximately linear relationship between %ABV and IBU that may have some curvature.  This positively correlated relationship is likely because people like drinking balanced beers.  Higher alcohol beers tend to also be maltier/sweeter (higher residual sugar) which balances the high bitterness (high IBUs).              

There appears to be a boundary near ABV=10% that most beers don't cross.  This may be due to limitations on product cost, beer yeast survival at higher ABV, or state/federal regulations on beer that is 10% ABV or more.     

The imputed IBU values are evident from the vertical bands at IBU=0, IBU=median(IBU), etc.


```{r ABV vs. IBU, echo=FALSE}

Beer %>% ggplot(aes(IBU,ABV)) + geom_count() + geom_smooth(method="lm") + ggtitle("Linear %ABV vs. IBU - Excludes Missing Data") + xlab("Bitterness (IBUs)") + ylab("Log(%ABV)")

Beer %>% ggplot(aes(IBU1,ABV1)) + geom_count() + geom_smooth(method="lm") + ggtitle("Linear %ABV vs. IBU - Predicted Values for Missing Data") + xlab("Bitterness (IBUs)") + ylab("Log(%ABV)")

#Exclude Missing Values
Beer %>% filter(!is.na(IBU) & !is.na(ABV)) %>% ggplot(aes(IBU,ABV*100)) + geom_count() + geom_smooth() + ggtitle("Non-Linear %ABV vs. IBU - Excludes Missing Values") + xlab("Bitterness (IBUs)") + ylab("%ABV") + coord_cartesian(xlim=c(0,150),ylim=c(0,12.6))

#With Imputed Missing Values
Beer %>% ggplot(aes(IBU1,ABV1*100)) + geom_count() + geom_smooth() + ggtitle("Non-Linear %ABV vs. IBU - Predicted Values for Missing Data") + xlab("Bitterness (IBUs)") + ylab("%ABV")  + coord_cartesian(xlim=c(0,150),ylim=c(0,12.6))

#Line Plot by Styles - Excluded NAs
#Beer %>% filter(!is.na(IBU) & !is.na(ABV)) %>% ggplot(aes(IBU,ABV*100,color=Style)) + geom_smooth(method="lm",se=F) + ggtitle("Scatterplot of Beer %ABV vs. IBU - Excludes Missing Values") + xlab("Bitterness (IBUs)") + ylab("%ABV") + theme(legend.position = "none")+ coord_cartesian(xlim=c(0,150),ylim=c(0,12.6))

#Line Plot by Styles - Predicted NAs
#Beer %>% ggplot() + geom_smooth(aes(IBU1,ABV1*100,color=Style),method="lm",se=F) + ggtitle("Scatterplot of Beer %ABV vs. IBU - Includes Predicted Values for Missing Data") + xlab("Bitterness (IBUs)") + ylab("%ABV") + theme(legend.position = "none") + coord_cartesian(xlim=c(0,150),ylim=c(0,12.6))
```

## 8.  Budweiser would also like to investigate the difference with respect to IBU and ABV between IPAs (India Pale Ales) and other types of Ale (any beer with "Ale" in its name other than IPA).  You decide to use KNN classification to investigate this relationship.  Provide statistical evidence one way or the other. You can of course assume your audience is comfortable with percentages . KNN is very easy to understand conceptually.

First, find beers with "IPA" or "Ale" directly in their name for the training set (ex: Ranger IPA).  Second, find IPAs and Ales without those identifiers in their names for the test set (ex: .   

```{r EDA IPAs vs. Ales, include=FALSE}
#How many beers have IPA or Ale in the beer's Name?

#Beers with "IPA" or "I.P.A." or "India Pale Ale" somewhere in the beer's Name
sum(str_detect(Beer$Name.Beer,"(IPA|I.P.A.|India Pale Ale)")) #382 IPAs
Beer$Name.Beer[str_detect(Beer$Name.Beer,"India")==T & str_detect(Beer$Name.Beer,"(IPA|India Pale Ale)")==F] #Other "India" beers to consider 

#Beers with "Ale" somewhere in the beer's Name, not including India Pale Ales
sum(str_detect(Beer$Name.Beer,"Ale")) - sum(str_detect(Beer$Name.Beer,"India Pale Ale")) #577 Ales 


### Not all IPAs or Ales have "IPA" or "Ale" in the beer's name ###


#How many beers have IPA or Ale as the beer's Style?

Style_ABV_df #Summary table of median ABV and Count by beer Style 
sum(str_detect(Style_ABV_df$Style,"IPA")) #5 IPA Styles
Style_ABV_df[str_detect(Style_ABV_df$Style,"IPA")==T,] #Summary of the 5 IPA Styles
IPA_count = sum(str_detect(Beer$Style,"IPA"))
IPA_count #571 IPAs total, includes IPAs that had missing ABV values 

sum(str_detect(Style_ABV_df$Style,"Ale")) #29 Ale Styles, including India Pale Ales
Style_ABV_df[str_detect(Style_ABV_df$Style,"Ale")==T,] #Summary of the Ale Styles
Ale_count = sum( str_detect(Beer$Style,"Ale") - str_detect(Beer$Style,"India Pale Ale") )
Ale_count #964 Ales total, not including IPAs


#What beers Styles aren't explicitly called "Ale"?

Ales = Beer %>% filter(str_detect(Style,"(IPA|Ale)")==F & str_detect(Name.Beer,"Ale")==T) %>% group_by(Style,Name.Beer) %>% summarize(meanABV=mean(ABV1))
Ales_df=as.data.frame(Ales)
Ales_df
dim(Ales_df) #66 beers that were missed, 1030 Ales total

```

### How well does kNN classify IPAs vs. Ales based on beer name alone?  Can it correctly classify IPAs and Ales that don't have "IPA" or "Ale" in the name?

Using predicted values for missing data, the kNN model predicts IPAs vs. Ales reasonably well with 90% accuracy.

There is a trade-off between Sensitivity and Specificity with the kNN model.  k=5 gives  higher sensitivity (classifies fewer Ales incorrectly as IPAs), but k=11 gives higher specificity (classifies fewer IPAs incorrectly as Ales).         

```{r kNN IPAs vs. Ales - Imputed, echo=FALSE}

#Filter the data to only keep Ales and IPAs
IPA_Ale = Beer %>% filter(str_detect(Style,"(IPA|Ale)")==T | str_detect(Name.Beer,"Ale")==T)
#dim(IPA_Ale) #1604 Ales and IPAs

#Label the beers as "IPA" or "Ale"
IPA_Ale$Class = as.factor( ifelse(str_detect(IPA_Ale$Style,"IPA")==T,"IPA","Ale") )
#summary(IPA_Ale$Class) 

#Plot the data - some overlap
#IPA_Ale %>% ggplot(aes(IBU1,ABV1,color=Class)) + geom_jitter()

#Training set based on beers with "Ale" or "IPA" in the Name
Train = IPA_Ale %>% filter(str_detect(IPA_Ale$Name.Beer,"(IPA|I.P.A.|Ale)")==T)
#dim(Train) #959 beers, 60% split
#summary(Train$Class)

#Plot the Training data 
Train %>% ggplot(aes(IBU1,ABV1,color=Class)) + geom_jitter() + ggtitle("Training Set")

#Test set based on beers without "Ale" or "IPA" in the Name (only in the Style)
Test = IPA_Ale %>% filter(str_detect(IPA_Ale$Name.Beer,"(IPA|I.P.A.|Ale)")==F)
#dim(Test) #645 beers, 40% split
#summary(Test$Class) #Count is more skewed towards Ales vs. Training Set

#Plot the Test data 
Test %>% ggplot(aes(IBU1,ABV1,color=Class)) + geom_jitter() + ggtitle("Test Set")

#kNN Modeling

#Find the optimal k 

CM_summary = data.frame(Accuracy = numeric(25), Sensitivity = numeric(25), 
                        Specificity = numeric(25), k = numeric(25))

for(i in 1:25)
{
  classifications = knn(Train[,c(16,13)], Test[,c(16,13)], Train$Class, prob = TRUE, k = i)
  table(Test$Class,classifications)
  CM = confusionMatrix(table(Test$Class,classifications))
  CM_summary$Accuracy[i] = CM$overall[1]
  CM_summary$Sensitivity[i]=CM$byClass[1]
  CM_summary$Specificity[i]=CM$byClass[2]
  CM_summary$k[i] = i
}

#Max Accuracy and Sensitivity flips between k=4 and k=5 due to random ties with k=4
#grep(max(CM_summary$Accuracy),CM_summary$Accuracy) #k=4 or 5
#grep(max(CM_summary$Sensitivity),CM_summary$Sensitivity) #k=4 or 5
#grep(max(CM_summary$Specificity),CM_summary$Specificity) #k=11
                     
CM_summary %>% ggplot() + geom_line(aes(k,Accuracy,color="Accuracy")) + geom_line(aes(k,Sensitivity,color="Sensitivity")) + geom_line(aes(k,Specificity,color="Specificity")) + geom_vline(xintercept=c(4,5,11)) + ylab("%") + ggtitle("Optimal k's for Classifying IPAs vs. other Ales")

```

### Now exclude the original missing values and repeat the kNN analysis.

In general, the model with excluded NA values performed worse than the model with predicted values for NA's.  

```{r kNN IPAs vs. Ales - Excluded, echo=FALSE}

IPA_Ale_NA = IPA_Ale %>% filter(!is.na(ABV) & !is.na(IBU)) 

#Plot the data - lots of overlap
IPA_Ale_NA %>% ggplot(aes(IBU,ABV,color=Class)) + geom_jitter()

#Training set based on beers with "Ale" or "IPA" in the Name
Train = IPA_Ale_NA %>% filter(str_detect(IPA_Ale_NA$Name.Beer,"(IPA|I.P.A.|Ale)")==T)
#dim(Train) #607 beers, ~60% split
#summary(Train$Class)

#Plot the Training data 
Train %>% ggplot(aes(IBU,ABV,color=Class)) + geom_jitter() + ggtitle("Training Set")

#Test set based on beers without "Ale" or "IPA" in the Name (only in the Style)
Test = IPA_Ale_NA %>% filter(str_detect(IPA_Ale_NA$Name.Beer,"(IPA|I.P.A.|Ale)")==F)
#dim(Test) #380 beers, ~40% split
#summary(Test$Class) #Count is more skewed towards Ales vs. Training Set

#Plot the Test data 
Test %>% ggplot(aes(IBU,ABV,color=Class)) + geom_jitter() + ggtitle("Test Set")

#kNN Modeling

#Find the optimal k 

CM_summary = data.frame(Accuracy = numeric(40), Sensitivity = numeric(40), 
                        Specificity = numeric(40), k = numeric(40))

for(i in 1:40)
{
  classifications = knn(Train[,c(16,13)], Test[,c(16,13)], Train$Class, prob = TRUE, k = i)
  table(Test$Class,classifications)
  CM = confusionMatrix(table(Test$Class,classifications))
  CM_summary$Accuracy[i] = CM$overall[1]
  CM_summary$Sensitivity[i]=CM$byClass[1]
  CM_summary$Specificity[i]=CM$byClass[2]
  CM_summary$k[i] = i
}


#grep(max(CM_summary$Accuracy),CM_summary$Accuracy) #k=5 or 37
#grep(max(CM_summary$Sensitivity),CM_summary$Sensitivity) #k=4
#grep(max(CM_summary$Specificity),CM_summary$Specificity) #k=37

CM_summary %>% ggplot() + geom_line(aes(k,Accuracy,color="Accuracy")) + geom_line(aes(k,Sensitivity,color="Sensitivity")) + geom_line(aes(k,Specificity,color="Specificity")) + geom_vline(xintercept=c(4,5,15,37)) + ylab("%") + ggtitle("Optimal k's for Classifying IPAs vs. other Ales")

```

## 9. Knock their socks off!  Find one other useful inference from the data that you feel Budweiser may be able to find value in.  You must convince them why it is important and back up your conviction with appropriate statistical evidence. 

### What makes a Pale Ale vs. IPA anyway?

IPAs are generally (but not always) hoppier and boozier than Pale Ales.  American Double / Imperial IPAs are pushing this trend to new limits.  Belgian Strong Pale Ales are outliers and may be more similar to Belgian Strong Ales rather than Pale Ales. 

```{r IPA vs. Pale Ale, echo=FALSE}

#Filter the data
IPA_PaleAle = IPA_Ale %>% filter(str_detect(Style,"(IPA|Pale Ale)")==T)
#summary(IPA_PaleAle$Style)
#dim(IPA_PaleAle)

#Label the beers as "American","English", or "Belgian"
IPA_PaleAle$Country = as.factor( ifelse(str_detect(IPA_PaleAle$Style,"American")==T,"American", ifelse(str_detect(IPA_PaleAle$Style,"English")==T,"English","Belgian")))
#summary(IPA_PaleAle$Country)     

#Categorize the beers by Style without Country
IPA_PaleAle$Class = gl(4,170,length=859,labels=c("Pale Ale","White IPA","Strong Pale Ale","IPA","Double / Imperial IPA"))
IPA_PaleAle$Class[grep("Ale",IPA_PaleAle$Style)]="Pale Ale"
IPA_PaleAle$Class[grep("IPA",IPA_PaleAle$Style)]="IPA"
IPA_PaleAle$Class[grep("White",IPA_PaleAle$Style)]="White IPA"
IPA_PaleAle$Class[grep("Double",IPA_PaleAle$Style)]="Double / Imperial IPA"
IPA_PaleAle$Class[grep("Strong",IPA_PaleAle$Style)]="Strong Pale Ale"
summary(IPA_PaleAle$Class)

IPA_PaleAle %>% ggplot() + geom_jitter(aes(IBU1,ABV1,color=Class)) + geom_smooth(aes(IBU1,ABV1),method='lm',color="black") + facet_wrap(~Country) + ggtitle("IPA vs. Pale Ale by Country - Predicted Missing Values")

IPA_PaleAle %>% ggplot() + geom_jitter(aes(IBU,ABV,color=Class)) + geom_smooth(aes(IBU,ABV),method='lm',color="black") + facet_wrap(~Country) + ggtitle("IPA vs. Pale Ale by Country - Excludes Missing Values")

IPA_PaleAle %>% ggplot(aes(IBU1,ABV1,color=Country)) + geom_jitter() + geom_smooth(method='lm',se=F) + ggtitle("IPA vs. Pale Ale by Country - Predicted Missing Values")

IPA_PaleAle %>% ggplot(aes(IBU,ABV,color=Country)) + geom_jitter() + geom_smooth(method='lm',se=F) + ggtitle("IPA vs. Pale Ale by Country - Excludes Missing Values")
```

### Do people actually like boozier beers?
Deeper dive on alcohol content and beer ratings from Kaggle data set.  Yes, people like boozier beers which is a statistically significant result using ANOVA.

```{r, ABV Reviews, echo=FALSE}
#Import Beer Advocate Review Data from Kaggle (https://www.kaggle.com/rdoume/beerreviews).  EDA of Overall Reviews and ABV.  There is some visual evidence to suggest that lower ratings have higher densities at low ABV values while we see the highest ratings have higher densities at higher ABV values.

p <- ggplot(beer_reviews, aes(factor(review_overall), beer_abv))
p + geom_violin(aes(fill = factor(review_overall))) + ggtitle("Violin Plot: ABV vs. Overall Review") + xlab("Overall Review") + ylab("ABV") + labs(fill = "Overall Review")


#Do the different overall rating groups have different ABV averages?
beer_reviews$group[beer_reviews$review_overall==1 | beer_reviews$review_overall==1.5]="1" 
beer_reviews$group[beer_reviews$review_overall==2| beer_reviews$review_overall==2.5]="2" 
beer_reviews$group[beer_reviews$review_overall==3 | beer_reviews$review_overall==3.5]="3" 
beer_reviews$group[beer_reviews$review_overall==4 | beer_reviews$review_overall==4.5]="4" 
beer_reviews$group[beer_reviews$review_overall==5]="5" 
#anova
model=aov(beer_reviews$beer_abv~beer_reviews$group)
summary(model)
TukeyHSD(model)
#The box plots clearly show an association with overall rating of a beer and the ABV of the beer.  As the rating increases it appears that the ABV also increases.
colors1 = c(rep("red",1),rep("orange",1),rep("yellow",1),rep("pink",1),rep("green",1))
boxplot(beer_reviews$beer_abv~beer_reviews$group,ylim=c(0,20), xlab = "Overall Rating", ylab = "ABV", main = "Boxplot of Ratings for ABV",col=colors1 )

#Correlation is significant at alpha 0.05
beer_reviews=na.omit(beer_reviews)
cor.test(beer_reviews$beer_abv,beer_reviews$review_overall)

```

## Conclusion

In conclusion, there is a wealth of information in this data set on Beers, Breweries, States, and beer characteristics (Style, IBU, ABV, etc.):  
- There are dominant brewing states in each region of the US.  
- Missing IBU data may be important for assessing emerging regional trends.
- Alcohol (ABV) content is shifting higher driven by demand for bigger, boozier beers.
- People generally like drinking beers that are balanced for alcohol/sweetness vs. bitterness.
- Some beer styles can be reliably identified based only on the IBU and ABV content.