---
title: "CaseStudy1"
author: "Anish Patel"
date: "2/22/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	comment = NULL
)
library(tidyverse)
library(maps)
library(mapproj)
```



```{r}
library(ggmap)
library(ggplot2)
library(ggmap)
library(maps)
library(mapdata)
library(openintro)
library(plotly)
library(billboarder)
library(vctrs)
library(dplyr)
states <- read.csv("~/Downloads/states.csv")
Breweries <- read.csv("~/Downloads/Breweries.csv")
Beers <- read.csv("~/Downloads/Beers.csv")
```


## Question 1 - How many Brewereries are in Each State?

```{r}

dat = Breweries

#obtain the count of the number of breweres by state arranged in decending order
state_count = dat %>% group_by(State) %>% count() %>% arrange(-n)


```

## Bar Chart of Brewery Count by State

```{r}

billboarder() %>%
  bb_barchart(state_count) %>%
  bb_labs(title = "Count of Breweries by State")
   
```

## Heatmap

```{r}
#Here we are creating heatmap to show quanties. 
heat_map = ggplot(data = states) + 
  geom_polygon(aes(x = long, y = lat, fill = n, group = group), color = "white") + 
  coord_fixed(1.3) +
  scale_fill_gradientn(colours = rev(heat.colors(10)),na.value = "grey90",
                       guide = guide_colourbar(barwidth = 5, barheight = 0.4,
                                               #put legend title on top of legend
                                               title.position = "top")) + 
  labs(title = "Count of Breweries by State (Continental US)")
  
ggplotly(heat_map)

  
```

## Question 2 - Merge data by Brew ID, and check first 6 and last 6 obs.

```{r}

#rename col to match Brewery Data set
colnames(Beers)[5]="Brew_ID"

#merge data by Brew_ID 
final=merge(Beers,Breweries,by=c("Brew_ID"))

head(final,n=6)
tail(final,n=6)

```

## Question 3 - How to deal with Missing Values
```{r}

#Of the missing values, 1005 are missing for "IBU", 5 missing values for "Style"", and 62 missing values for ABV, and 1,005 missing values for IBU.  


#One method would be to omit the observations with missing data which isn't ideal since we will lose quite a few observations

omita=na.omit(final, cols = "ABV")
omiti=na.omit(final, cols = "IBU")

#The other method would be to impute the median values of the variable to the missing variables.

#Find Medians of Style, IBU, and ABV

```

##Question 4 - What is the median alcohol content and international bitterness unit for each state?

```{r}

#find median of ABV and IBU
 omita%>%group_by(State)%>%summarise(median(ABV),median(IBU))

```

## Question 5 - Which state has the most bitter beer and highest ABV

When we exclude all of the missing values, we find that Maine has both the highest median of ABV and IBU amongst all states.  THe highest ABV is London Balling from KY.  The most bitter beer (highest IBU) is "Bitter Bitch Imperial IPA" from OR.

```{r}

medianalc=as.data.frame(omita%>%group_by(State)%>%summarise(median(ABV)))
medianibu=as.data.frame(omiti%>%group_by(State)%>%summarise(median(IBU)))

                        
#label columns
colnames(medianalc)=c("State","ABV")
colnames(medianibu)=c("State","IBU")

#sort in descending order
medianalc = medianalc %>% arrange(-ABV)
medianibu = medianibu %>% arrange(-IBU)

#create charts

billboarder() %>%
  bb_barchart(medianalc) %>%
  bb_labs(title = "Median ABV of Breweries by State")

billboarder() %>%
  bb_barchart(medianibu) %>%
  bb_labs(title = "Median IBU of Breweries by State")

#Find beer with MAX ABV and MAX IBU
omita[omita$ABV==max(omita$ABV),] 
omiti[omiti$IBU==max(omiti$IBU),] 


  
```
## Question 6 - Comment on Distribution of ABV and summary statistics
We see that the distribution to be right skewed (the mean of 0.05977 is greater than the median of 0.05600) as seen in the histogram.  We wee a range of 0.001 - 0.128 in the data for ABV.  There are 62 missing values that we exclude from the analysis.
```{r}


summary(omita$ABV)
final %>% ggplot(aes(ABV*100)) + geom_histogram() + xlab("Alcohol by Volume") + ggtitle("Distribution of Beer ABV") 


```

## Question 7 - Is there a relationship between IBU and ABV?
Looking at the plot, there does appear to be an association between IBU and ABV.  As the ABV increases, we see IBU tends to increase as well.  Theres seems to be a direct positive relationship between the two variables.  There are some extreme outliers with extremely high ABV seen in the far right of the plot.
```{r}

ggplot(data=omita) + geom_point(aes(x=ABV,y=IBU)) + ggtitle ("IBU vs ABV (Excluding all Missing Values)")

```

